---
title: "Minimum Viable Demo Setup Script"
author: "Carl Howe"
date: "6/22/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

This document is used to generate some data for analysis during the RStudio Connect Minimum Viable Demo. I mean it to be run prior to performing the demo so that all the data necessary for the demo is in place.

The concept is
that a company called Digitcomputronimatics has 3 products: WidgetBasic (represented by variable widget_basic), WidgetMaster (widget_master), and WidgetKillDozer (widget_killdozer). Sales overall are increasing. However, the sales of the WidgetBasic and WidgetKillDozer are masking a decline in sales for the WidgetMaster, which our analysis will reveal.

```{r data generation, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tibbletime)
library(lubridate)
set.seed <- 1
number_of_days <- 366+365+180
sales_period <- lubridate::ymd(make_datetime(year=2016, day=1:(number_of_days)))

trend <- function(x, amount) {
  x * amount
}

seasonal <- function(x, offset, amount) {
  amount * sin(2 * pi * (x + offset)/ 365)
}

noise <- function(x, amount) {
  runif(n = x, min = -amount, max = amount)
}

widget_sales <- tibble('date' = sales_period, 
                  'widget_basic_sales'     =  ts(500 + 
                                              trend(1:number_of_days, 150/365) + 
                                              seasonal(1:number_of_days, 270, 30) + 
                                              noise(1:number_of_days, 300), frequency=365),
                  'widget_killdozer_sales' = ts(500 +
                                             trend(1:number_of_days, 75/365) + 
                                             seasonal(1:number_of_days, 315, 30) + 
                                             noise(1:number_of_days, 200), frequency=365),
                  'widget_master_sales'    = ts(600 + 
                                             trend(1:number_of_days, -75/365) + 
                                             seasonal(1:number_of_days, 0, 60) + 
                                             noise(1:number_of_days, 400), frequency=365))

widget_sales <- widget_sales %>% 
                       mutate(total_sales = widget_basic_sales + widget_killdozer_sales + widget_master_sales)
widget_sales <- as_tbl_time(widget_sales, index=date)
print(widget_sales %>% filter_time(~ '2018'))
widget_sales_gathered <- gather(widget_sales, key='product', value="sales", 
                                widget_basic_sales, widget_killdozer_sales, widget_master_sales, total_sales) %>%
                         arrange(date)
## ggplot(data=widget_sales, aes(x=date, y=total_sales)) + geom_line() + scale_x_datetime() + scale_y_continuous()
ggplot(data=widget_sales_gathered, aes(x=date, y=sales, color=product, group=product)) + 
    geom_line(alpha=0.3) + geom_smooth(span=0.3, method="loess", se=FALSE) + scale_x_date() + scale_y_continuous() +
    ggtitle("Daily Widget Sales Data")
system("mkdir /tmp/mvd-data")
write_csv(widget_sales, "/tmp/mvd-data/widget_sales_daily.csv")
```


Generate weekly and monthly versions of the data for the use of the demo and write them out into the file system. Normally, we do this in a database in a real world application, but this is just a minimal demo.

```{r collapsing data, echo=FALSE, message=FALSE, warning=FALSE}
widget_sales_weekly <- widget_sales %>% collapse_by("weekly") %>% dplyr::group_by(date) %>% dplyr::summarise_all(sum)
widget_sales_weekly_gathered <- gather(widget_sales_weekly, key='product', value="sales", 
                                widget_basic_sales, widget_killdozer_sales, widget_master_sales, total_sales) %>%
                         arrange(date)
widget_sales_monthly <- widget_sales %>% collapse_by("monthly") %>% dplyr::group_by(date) %>% dplyr::summarise_all(sum)
widget_sales_monthly_gathered <- gather(widget_sales_monthly, key='product', value="sales", 
                                widget_basic_sales, widget_killdozer_sales, widget_master_sales, total_sales) %>%
                         arrange(date)
write_csv(widget_sales_weekly, "/tmp/mvd-data/widget_sales_weekly.csv")
ggplot(data=widget_sales_weekly_gathered, aes(x=date, y=sales, color=product, group=product)) + 
    geom_line() + scale_x_date() + scale_y_continuous() +
    ggtitle("Weekly Widget Sales Data")
ggplot(data=widget_sales_monthly_gathered, aes(x=date, y=sales, color=product, group=product)) + 
    geom_line() + scale_x_date() + scale_y_continuous() +
    ggtitle("Monthly Widget Sales Data")
write_csv(widget_sales_monthly, "/tmp/mvd-data/widget_sales_monthly.csv")
```

Assuming all the data wrote out properly, you're now ready to begin the demo.
